---
title: "Generic Workflow Template" 
author: "Author: FirstName LastName"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`" 
output:
  BiocStyle::html_document:
    toc_float: true
    code_folding: show
  BiocStyle::pdf_document: default
package: systemPipeR
vignette: |
  %\VignetteIndexEntry{Basic Generic Template}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
fontsize: 14pt
bibliography: bibtex.bib
---

```{css, echo=FALSE}
pre code {
white-space: pre !important;
overflow-x: scroll !important;
word-break: keep-all !important;
word-wrap: initial !important;
}
```

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=60, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")), 
    tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

# Workflow environment

This is a _Generic_ template for a systemPipeR workflow provided by
[systemPipeRdata](https://bioconductor.org/packages/devel/data/experiment/html/systemPipeRdata.html), 
a companion package to [systemPipeR](https://www.bioconductor.org/packages/devel/bioc/html/systemPipeR.html). 
Like other workflow templates, it can be loaded with a single command. Once loaded, users have the
flexibility to utilize the templates as they are or modify them as needed. More
in-depth information can be found in the main vignette of [systemPipeRdata](https://bioconductor.org/packages/devel/data/experiment/vignettes/systemPipeRdata/inst/doc/systemPipeRdata.html). The
`Generic` template presented here is special that it provides a workflow
skelleton intended to be used as a starting point for building new workflows.
Basic workflow steps are included to illustrate how to design command-line (CL)
and R-based workflow steps, as well as R Markdown code chunks that are not part
of a workflow. The details about contructing workflow steps are explained in 
`systemPipeR's` main [vignette](https://www.bioconductor.org/packages/devel/bioc/vignettes/systemPipeR/inst/doc/systemPipeR.html#52_Constructing_workflows).

The following illustrates how to load the basic workflow steps into 
a `SAL` workflow container (`SYSargsList`) either one-by-one in interactive mode 
or with a single load command, and then run the workflow with a single command. 

This `Generic` workflow template includes the following four steps.

1. R step: export tabular data to files 
2. CL step: compress files
3. CL step: uncompress files 
4. R step: import files and plot summary statistics

##  Initialize workflow environment

The `SPRproject` function initializes a new workflow environment. This function
call creates an empty `SAL` workflow container instance and at the same time a
linked project log directory that acts as a flat-file database of a workflow.
For additional details, please visit this
[section](https://www.bioconductor.org/packages/devel/bioc/vignettes/systemPipeR/inst/doc/systemPipeR.html#5_Detailed_tutorial)
in `systemPipeR's` main vignette.  

```{r create_workflow, message=FALSE, eval=FALSE}
library(systemPipeR)
sal <- SPRproject()
sal
```

## Load packages 

Next, the `systemPipeR` package needs to be loaded in a workflow.

```{r load_library, eval=FALSE, spr=TRUE}
appendStep(sal) <- LineWise(
    code = {
    library(systemPipeR)
    }, 
    step_name = "load_library"
)
```

## Step 1: Export tabular data to files

This is the first data processing step. In this case it is an R step that uses the `LineWise` 
function to define the workflow step, and appends it to the `SAL` workflow container.

```{r export_iris, eval=FALSE, spr=TRUE}
appendStep(sal) <- LineWise(code={
    mapply(
      FUN = function(x, y) write.csv(x, y),
      x = split(iris, factor(iris$Species)),
      y = file.path("results", paste0(names(split(iris, factor(iris$Species))), ".csv"))
    )
    }, 
  step_name = "export_iris", 
  dependency = "load_library"
)
```

## Step 2: Compress data

The following adds a CL step that uses the `gzip` software to compress the files that were 
generated in the previous step.

```{r gzip, eval=FALSE, spr=TRUE, spr.dep=TRUE}
targetspath <- system.file("extdata/cwl/gunzip", "targets_gunzip.txt", package = "systemPipeR")
appendStep(sal) <- SYSargsList(
    step_name = "gzip", 
    targets = targetspath, dir = TRUE,
    wf_file = "gunzip/workflow_gzip.cwl", input_file = "gunzip/gzip.yml",
    dir_path = "param/cwl",
    inputvars = c(FileName = "_FILE_PATH_", SampleName = "_SampleName_"), 
    dependency = "export_iris"
)
```

## Step 3: Uncompress data

Next, the output files (here compressed `gz` files), that were generated by the
previous `gzip` step, will be uncompressed in the current step with the `gunzip`
software. 

```{r gunzip, eval=FALSE, spr=TRUE}
appendStep(sal) <- SYSargsList(
    step_name = "gunzip", 
    targets = "gzip", dir = TRUE,
    wf_file = "gunzip/workflow_gunzip.cwl", input_file = "gunzip/gunzip.yml",
    dir_path = "param/cwl",
    inputvars = c(gzip_file = "_FILE_PATH_", SampleName = "_SampleName_"), 
    rm_targets_col = "FileName", 
    dependency = "gzip"
)
```

## Step 4: Import tabular files and visualize data

Imports the tabular files from the previous step back into R, performs some summary 
statistics and plots the results as bar diagrams.

```{r stats, eval=FALSE, spr=TRUE}
appendStep(sal) <- LineWise(code={
    # combine all files into one data frame
    df <- lapply(getColumn(sal, step="gunzip", 'outfiles'), function(x) read.delim(x, sep=",")[-1])
    df <- do.call(rbind, df)
    # calculate mean and sd for each species
    stats <- data.frame(cbind(mean=apply(df[,1:4], 2, mean), sd=apply(df[,1:4], 2, sd)))
    stats$species <- rownames(stats)
    # plot
    plot <- ggplot2::ggplot(stats, ggplot2::aes(x=species, y=mean, fill=species)) + 
        ggplot2::geom_bar(stat = "identity", color="black", position=ggplot2::position_dodge()) +
        ggplot2::geom_errorbar(
            ggplot2::aes(ymin=mean-sd, ymax=mean+sd), 
            width=.2,
            position=ggplot2::position_dodge(.9)
        )
    plot
    }, 
    step_name = "stats", 
    dependency = "gunzip", 
    run_step = "optional"
)
```

## Version Information

```{r sessionInfo, eval=FALSE, spr=TRUE}
appendStep(sal) <- LineWise(
    code = {
    sessionInfo()
    }, 
    step_name = "sessionInfo", 
    dependency = "stats")
```


# Manage the workflow

## Interactive job submissions in a single machine

For running the workflow, `runWF` function will execute all the steps store in 
the workflow container. The execution will be on a single machine without 
submitting to a queuing system of a computer cluster. 

```{r runWF, eval=FALSE}
sal <- runWF(sal)
```

## Visualize workflow

_`systemPipeR`_ workflows instances can be visualized with the `plotWF` function.

```{r plotWF, eval=FALSE}
plotWF(sal, rstudio = TRUE)
```

## Checking workflow status

To check the summary of the workflow, we can use:

```{r statusWF, eval=FALSE}
sal
statusWF(sal)
```

## Accessing logs report

_`systemPipeR`_ compiles all the workflow execution logs in one central location,
making it easier to check any standard output (`stdout`) or standard error
(`stderr`) for any command-line tools used on the workflow or the R code stdout.

```{r logsWF, eval=FALSE}
sal <- renderLogs(sal)
```

# About the workflow
## Tools used 
To check command-line tools used in this workflow, use `listCmdTools`, and use `listCmdModules`
to check if you have a modular system.

The following code will print out tools required in your custom SPR project in the report. 
In case you are running the workflow for the first time time and do not have a project yet, or you 
just want to browser this workflow, following code displays the tools required by default.
```{r list_tools}
if(file.exists(file.path(".SPRproject", "SYSargsList.yml"))) {
    local({
        sal <- systemPipeR::SPRproject(resume = TRUE)
        systemPipeR::listCmdTools(sal)
        systemPipeR::listCmdModules(sal)
    })
} else {
    cat(crayon::blue$bold("Tools and modules required by this workflow are:\n"))
    cat(c("gzip", "gunzip"), sep = "\n")
}
```

## Session Info
This is the session information for rendering this report. To access the session information
of workflow running, check HTML report of `renderLogs`. 
```{r report_session_info, eval=TRUE}
sessionInfo()
```

# References
